# Common Graphics {#common_graphics}

```{r cg_setup, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
```


In this section we will show the R code used to generate some common statistical graphics. The graphics will be based on built-in R datasets so you can test them easily and then change the dataset and variables (column headings) parts of the code to easily plot your own data.

## Barchart

Barcharts are sometime sused to plot numerical data, including counts,  for a set of categories. For our first example we'll use the `mpg` dataset. This lists `nrow(mpg)` cars and includesd data on their manufacturer and fuel efficiency. We can look at the top of the dataset with these commands...

```{r cg_mpg}
print(mpg, width = Inf)
```

To plot a barchart showing the number of cars in the dataset from each manufactirer we can use the `ggplot()` function with manufacturer as the `x` aesthetic and using the `geom_bar` geom.

```{r cg_mpg_barchart}
ggplot(mpg, aes(x = manufacturer)) +
  geom_bar()
```


## Dotchart
## Histogram
## Frequency polygram
## Scatterplot
## Scatterplot with smoother
